### 2017-05-31 doc/intro.tex
Models and intro

As our case study we picked the {\sc Imaginet} model introduced by \label{explainimaginet}
\namecite{chrupala2015learning}. It is a multi-task, multi-modal
architecture consisting of two Gated-Recurrent Unit (GRU)
\cite{cho2014properties,chung2014empirical} pathways and
a shared word embedding matrix. One of the GRUs ({\sc Visual}) is
trained to predict image vectors given image descriptions, while the other
pathway ({\sc Textual}) is a language model. This particular
architecture allows a comparative analysis of the hidden activations
patterns between networks trained on two very different tasks, while
keeping the training data and the word embeddings fixed. Recurrent neural
language models akin to {\sc Textual} which are trained to predict the
next symbol in a sequence are relatively well understood, and there
have been some attempts to analyze their internal states e.g.
\cite{elman1991distributed,karpathy2015visualizing}. In
constrast, {\sc Visual} maps a complete sequence of words to
a representation of a corresponding visual scene and is a less
commonly encountered, but a more interesting model from the point of
view of representing meaning conveyed via linguistic structure.

\todo[inline]{once we finalise our experiments, we should update the following paragraphs and get rid of the macro/micro distinction.}
We report a thorough \emph{macro level} quantitative analysis to
provide a linguistic interpretation of the networks' activation
patterns. These experiments rely on a novel method we
call \emph{omission score} to measure the importance of input tokens 
to the final prediction of models that compute distributed representations 
of sentences. This can be applied to various RNN architectures, Recursive Neural Networks 
and Convolutional Neural Networks.\label{edit:introomissiongeneral}
Based on these omission scores our experiments show that the image prediction 
pathway in general pays special attention to words of syntactic categories that 
carry semantic content. Furthermore, we observe that on top of
lexical features it learns important aspects of the information structure of sentences,
and treats the same input tokens differently depending on their
grammatical function in the sentence. In contrast, the language model
is more sensitive to the local syntactic characteristics of the input
sentences.

>>>>

As our case study we picked the {\sc Imaginet} model introduced by \label{explainimaginet}
\namecite{chrupala2015learning}. It is a multi-task, multi-modal
architecture consisting of two Gated-Recurrent Unit (GRU)
\cite{cho2014properties,chung2014empirical} pathways and
a shared word embedding matrix. One of the GRUs ({\sc Visual}) is
trained to predict image vectors given image descriptions, while the other
pathway ({\sc Textual}) is a language model, trained to sequentially predict each
word in the descriptions. This particular
architecture allows a comparative analysis of the hidden activations
patterns between networks trained on two different tasks, while
keeping the training data and the word embeddings fixed. Recurrent neural
language models akin to {\sc Textual} which are trained to predict the
next symbol in a sequence are relatively well understood, and there
have been some attempts to analyze their internal states e.g.
\cite{elman1991distributed,karpathy2015visualizing}. In
constrast, {\sc Visual} maps a complete sequence of words to
a representation of a corresponding visual scene and is a less
commonly encountered, but a more interesting model from the point of
view of representing meaning conveyed via linguistic structure.
For comparison, we also consider a standard standalone language model.

\todo[inline]{once we finalise our experiments, we should update the following paragraphs and get rid of the macro/micro distinction.}
We report a thorough \emph{macro level} quantitative analysis to
provide a linguistic interpretation of the networks' activation
patterns. These experiments rely on a novel method we
call \emph{omission score} to measure the importance of input tokens 
to the final prediction of models that compute distributed representations 
of sentences. This can be applied to various RNN architectures, Recursive Neural Networks 
and Convolutional Neural Networks.\label{edit:introomissiongeneral}

Based on these omission scores our experiments show that the image
prediction pathway in general pays special attention to syntactic
categories which carry semantic content, and especially nouns. More
surprisingly, this pathway also learns learns to treat word types
differently depending on their grammatical function and their position
in the sequential structure of the sentence. In contrast, the textual
pathway and the standalone language model are especially sensitive to
the local syntactic characteristics of the input sentences.

### 2017-05-31 doc/models.tex
Models and intro

Formally, each sentence is mapped to two sequences of hidden states, 
one by {\sc Visual} and the other by {\sc Textual}:
\vspace{-.2cm}
\begin{align}
  \mathbf{h}^{V}_t = \mathrm{GRU}^V(\mathbf{h}^{V}_{t-1},\mathbf{x}_t)\\
  \mathbf{h}^{T}_t = \mathrm{GRU}^T(\mathbf{h}^{T}_{t-1},\mathbf{x}_t)
\vspace{-.1cm}
\end{align}
%
At each time step {\sc Textual} predicts the next word in the sentence
$S$ from its current hidden state $\mathbf{h}^{T}_t$, while {\sc
  Visual} predicts the image-vector\footnote{Representing the full image, 
  extracted from the pre-trained Convolutional Neural Network of 
  \namecite{simonyan2014very}. \label{edit:dumdumeddy}}
$\hat{\mathbf{i}}$ from its last
hidden representation $\mathbf{h}^{V}_t$.
%
\vspace{-.2cm}
\begin{align}
   \hat{\mathbf{i}} &= \mathbf{V} \mathbf{h}^{V}_\tau \\
    p(S_{t+1}|S_{1:t}) &= \mathrm{softmax}(\mathbf{L}\mathbf{h}^T_t)
\vspace{-.1cm}
\end{align}
%
The loss function is a multi-task objective which penalizes error on
the visual and the textual targets simultaneously. The objective
combines cross-entropy loss $L^{T}$ for the word predictions
and cosine distance $L^V$ for the image
predictions.
%
%\vspace{-.2cm}
\begin{align}
%\begin{equation}
%\label{eq:lossce}
&L^{T}(\theta) = {-} \frac{1}{\tau}\sum_{t=1}^\tau \log p(S_t|S_{1:t}) \\
%\end{equation}
%
%\vspace{-.2cm}
%\begin{equation}
%\label{eq:losscos}
&L^V(\theta) =  1 - \frac{\hat{\mathbf{i}} \cdot \mathbf{i}}{\| \hat{\mathbf{i}} \| \| \mathbf{i} \|} \\
%\end{equation}
%
%\vspace{-.2cm}
%\begin{equation}
%\label{eq:losscombo}
&L = \alpha L^T + (1-\alpha)L^{V}
\end{align}
%L^V(\theta) &= \frac{1}{K}\sum_{k=1}^K(\hat{i}_k-i_k)^2
%

\noindent
For more details about the {\sc Imaginet} model and 
its performance see \namecite{chrupala2015learning}.

In our version of the model we add two minor modifications 
compared to the original {\sc Imaginet}. As the performance of
{\sc Visual} is measured on image-retrieval which is based on cosine 
distances, we use cosine distance as the visual loss.
Moreover, we observe that using standardized image vectors, where each dimension 
is transformed by subtracting the mean and dividing by standard deviation, further
improves performance.

>>>>

Formally, each sentence is mapped to two sequences of hidden states, 
one by {\sc Visual} and the other by {\sc Textual}:
\vspace{-.2cm}
\begin{align}
  \mathbf{h}^{V}_t = \mathrm{GRU}^V(\mathbf{h}^{V}_{t-1},\mathbf{x}_t)\\
  \mathbf{h}^{T}_t = \mathrm{GRU}^T(\mathbf{h}^{T}_{t-1},\mathbf{x}_t)
\vspace{-.1cm}
\end{align}
%
At each time step {\sc Textual} predicts the next word in the sentence
$S$ from its current hidden state $\mathbf{h}^{T}_t$, while {\sc
  Visual} predicts the image-vector\footnote{Representing the full image, 
  extracted from the pre-trained Convolutional Neural Network of 
  \namecite{simonyan2014very}. \label{edit:dumdumeddy}}
$\hat{\mathbf{i}}$ from its last
hidden representation $\mathbf{h}^{V}_t$.
%
\vspace{-.2cm}
\begin{align}
   \hat{\mathbf{i}} &= \mathbf{V} \mathbf{h}^{V}_\tau \\
    p(S_{t+1}|S_{1:t}) &= \mathrm{softmax}(\mathbf{L}\mathbf{h}^T_t)
\vspace{-.1cm}
\end{align}
%
The loss function is a multi-task objective which penalizes error on
the visual and the textual targets simultaneously. The objective
combines cross-entropy loss $L^{T}$ for the word predictions
and cosine distance $L^V$ for the image
predictions\footnote{Note that the original formulation in
  \namecite{chrupala2015learning} uses mean squared error instead; as the performance of
{\sc Visual} is measured on image-retrieval which is based on cosine 
distances, we use cosine distance as the visual loss here.}, weighting them with the parameter $\alpha$ (set to $0.1$).
%
%\vspace{-.2cm}
\begin{align}
%\begin{equation}
%\label{eq:lossce}
&L^{T}(\theta) = {-} \frac{1}{\tau}\sum_{t=1}^\tau \log p(S_t|S_{1:t}) \\
%\end{equation}
%
%\vspace{-.2cm}
%\begin{equation}
%\label{eq:losscos}
&L^V(\theta) =  1 - \frac{\hat{\mathbf{i}} \cdot \mathbf{i}}{\| \hat{\mathbf{i}} \| \| \mathbf{i} \|} \\
%\end{equation}
%
%\vspace{-.2cm}
%\begin{equation}
%\label{eq:losscombo}
&L = \alpha L^T + (1-\alpha)L^{V}
\end{align}
%L^V(\theta) &= \frac{1}{K}\sum_{k=1}^K(\hat{i}_k-i_k)^2
%

\noindent
For more details about the {\sc Imaginet} model and its performance
see \namecite{chrupala2015learning}. Note that we introduce a small change in
the image representation: we observe that using standardized image
vectors, where each dimension is transformed by subtracting the mean
and dividing by standard deviation, improves performance.
